---
title: "Statistics"
author: "Chun-Hui Gao"
date: "`r Sys.Date()`"
output: "html_document"
---

# Loading library

Statistics were performed maily with the `stats` and `rstatix` package in R version 3.5.3. First of all, we loaded the packages and the `tidyverse` package.

```{r data, message=F}
library(tidyverse)
library(stats)
library(rstatix)

# order of NPs
np_levels <- c("CK","Fe2O3", "ZnO", "CeO2", "Fe3O4", "Al2O3", "CuO", "TiO2")
```


# Figure 3: Comparison of Ct values

This part is for the comparision of Ct values. Ct values for each RT-PCR reaction was stored in "RT-PCR_results.csv". The experiment has repeat four times, so each group has 4 Ct values.

The following code block read in the result.

```{r compaision of Ct values, message=F}
results <- read_csv("data/RT-PCR_results.csv")
results$nps <- factor(results$nps, levels = np_levels)
```

Analysis then can be performed for Ex-taq or Phusion separately.

## Fig. 3A Ex-taq.

As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the groups highlighted with “**” in the model summary.

```{r}
data <- filter(results,enzyme=="extaq")
aov <- aov(CT~nps,data=data)
summary(aov)
```

As the ANOVA test is significant, we can compute Tukey HSD (**Tukey Honest Significant Differences**, R function: `TukeyHSD()`) for performing multiple pairwise-comparison between the means of groups.

```{r}
TukeyHSD(aov)
```

It can be seen from the output, that the difference between CK and Al2O3 is significant with an adjusted p-value of 0.046.


The ANOVA test assumes that, the data are normally distributed and the variance across groups are homogeneous. It’s possible to use **Levene’s test** to check the homogeneity of variances. The function `leveneTest()` [in `car` package] will be used:



```{r}
car::leveneTest(CT~nps,data=data)
```

From the output above we can see that the p-value is not less than the significance level of 0.05. This means that there is no evidence to suggest that the variance across groups is statistically significantly different. Therefore, we can assume the homogeneity of variances in the different NP groups.

Normality assumption can be checked by the **Shapiro-Wilk test** on the ANOVA residuals. 

```{r}
shapiro.test(x=residuals(aov))
```
The output above finds no indication that normality is violated (W = 0.97, p = 0.75). We can assume normality.

## Fig. 3B Phusion

Normality assumption can be checked by the **Shapiro-Wilk test** on the ANOVA residuals as well. 


```{r}
data <- filter(results,enzyme=="phusion")
aov <- aov(CT~nps,data=data)
shapiro.test(x=residuals(aov))
```
The output above finds that normality is violated (W = 0.81, p = 0.001). We can not assume normality.

Therefore, we have to use **Kruskal-Wallis rank sum test** for this data, as this test can be used when ANOVA assumptions are not met.

```{r}
kruskal.test(CT~nps,data=data)
```

As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between differnt NP treatments.


From the output of the **Kruskal-Wallis test**, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different.

It’s possible to use the function `pairwise.wilcox.test()` to calculate pairwise comparisons between group levels with corrections for multiple testing.

```{r}
pairwise.wilcox.test(data$CT,data$nps,p.adjust.method = "BH")
```

The pairwise comparison shows that, CK and four NPs conditions are all significant different (p < 0.05).

# Figure 4

```{r compaision of overall error frequences, message=F}
# error rate and snp frequency were estimated using amplicon sequencing and mothur software
error_rate <- read_csv("data/error_rate.csv")

# read meta data of sequencing library
meta <- read_tsv("data/meta.txt")
meta$nps <- factor(meta$nps, levels = np_levels)

# join data
error_rate <- left_join(error_rate,meta)
```

The analysis is similar to that has been mentioned on the above.

## Fig. 4A Ex-taq

```{r}
data <- error_rate %>% filter(enzyme=="extaq")
aov <- aov(error_rate~nps,data)
summary(aov)
```

As the p-value is greater than the significance level 0.05, we can conclude that there is no significant differences of error rate between differnt NP treatments.

```{r}
car::leveneTest(error_rate~nps,data=data)
shapiro.test(x=residuals(aov))
```

The output above finds ANOVA assumptions is valid, therefore, the test result is confidential.

## Fig. 4B Phusion

```{r}
data <- error_rate %>% filter(enzyme=="phusion")
aov <- aov(error_rate~nps,data)
summary(aov)
```

As the p-value is greater than the significance level 0.05, we can conclude that there is no significant differences of error rate between differnt NP treatments.

```{r}
car::leveneTest(error_rate~nps,data=data)
shapiro.test(x=residuals(aov))
```

The output above finds ANOVA assumptions is validty, therefore, the test result is confidential.


# Figure 5

```{r compaision of SNV frequences, message=F}
snp_freq <- read_csv("data/snp_freq.csv")
snp_freq <- left_join(snp_freq,meta)
```
## Fig. 5A Ex-taq

```{r}
data <- snp_freq %>% filter(enzyme=="extaq")
data %>% group_by(snv) %>% 
  anova_test(freq~nps) %>%
  mutate(p.adj=p.adjust(p,method = "BH"))
```

Since multiple comparisions were conducted to different groups, we need to adjust the p-values. Here, the **Benjamini & Hochberg ("BH") adjustment method** was applied.

As the `p.adj` is all greater than the significance level 0.05, we can conclude that there is no significant differences of SNV frequency between different NP treatments.


## Fig. 5B Phusion

```{r}
data <- snp_freq %>% filter(enzyme=="phusion")
data %>% group_by(snv) %>% 
  anova_test(freq~nps) %>%
  mutate(p.adj=p.adjust(p,method = "BH"))
```

As the p.adj is greater than the significance level 0.05, we can conclude that there is no significant differences of SNV frequency between differnt NP treatments.

# Session Info

```{r}
sessionInfo()
```

